Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-10 docs/ACM/paper.tex
Adding conclusion to Sec. 5.

\begin{figure}[t!]
\begin{small}
\begin{eqnarray*}
\begin{tabular}{lc}
\textsc{[Instr]} &
W(\mbox{instr}) = \mbox{given by the spec.}
\\\\
\textsc{[Seq]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2}{W(S_1;S_2;) = w_1 + w_2}
\\\\
\textsc{[Branch]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ W(S_3) = w_3}
{W(\mbox{if}(S_1) \ S_2; \ \mbox{else} \ S_3;) = w_1 + w_2 + w_3}
\\\\
\textsc{[LoopInf]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ \mathit{Iter}(S_1) = \infty}
{W(\mbox{while} (S_1) \ S_2;) = 10 \times (w_1 + w_2)}
\\\\
\textsc{[LoopExp]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ \mathit{Iter}(S_1) = E}
{W(\mbox{while} (S_1) \ S_2;) = E \times (w_1 + w_2)}
\end{tabular}
\end{eqnarray*}
\end{small}
\caption{\label{fig:swe}Estimating the cost of tasks.}
\end{figure}

@@

Similarly, recursion bounding (Sec.~\ref{sub:ir}), although
simple, is effective in eliminating the excessive number of tasks.
Figure~\ref{fig:Optimizations} shows the effect of this optimization upon
three \textsf{BSC-Bots} benchmarks: \textsf{health}, \textsf{nqueens} and
\textsf{strassen}.
These programs were designed to illustrate the parallelization of
divide-and-conquer algorithms~\cite{Duran09}, and contain a large number of
recursive calls (see \textbf{RFC}) in Figure~\ref{fig:Optimizations}.
Pruning at higher levels of the recursion tree lets them deliver
non-negligible speedups onto the baseline programs.
Were pruning absent, then we would observe slowdowns in \textsf{health}
and \textsf{nqueens}.

@@

The benchmarks used in Section~\ref{sub:performance} have been written to
demonstrate the power of different parallel systems; hence, they have already
been written in a way that simplifies the discovery of parallelism in programs.
In this section, we address the following question: ``can \Taskminer" find
parallelism in general programs?
To answer it, we have applied \Taskminer{} onto the 219 C programs available in
the LLVM test suite, and have compared the annotated version
against the sequential version compiled with \textsf{clang} -O3.
\Taskminer{} annotates a benchmark if it can find symbolic bounds to every
memory access (load or store) used within a region marked as a task candidate.
Under this constraint, we have been able to discover tasks in 63 benchmarks.
Figure~\ref{fig:TM_Versatility} relates the number of tasks and the number of
instructions in benchmarks.
To avoid counting multiple C files for the same benchmark,
Figure~\ref{fig:TM_Versatility} contains only benchmarks that are
constituted by a single file (present in LLVM's \textsf{SingleSource} folder).
As this experiment demonstrates, \Taskminer{} {\em can annotate a non-trivial
number of real-world benchmarks}.

@@

Most of the benchmarks, e.g., 27, that have been automatically annotated did not
give us speedups, because the regions marked as tasks were too small to influence
the program's runtime.
We have observed slowdowns in 17 benchmarks.
In this case, interactions between data-structures end up forcing the OpenMP
runtime to serialize execution of tasks.
And we have measured speedups above 5\% in 19 benchmarks.
In one case, \textsf{Misc/lowercase}, the speedup is above 3.5x, and in
three cases, it goes beyond 1.5x.
Figure~\ref{fig:TM_GeneralProgs} summarizes these findings.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_GeneralProgs}
\caption{Speedups obtained by \Taskminer{} when applied onto the LLVM
test suite.
The larger the bar, the better.
\textbf{LoC} stands for ``Lines of C".}
\label{fig:TM_GeneralProgs}
\end{center}
\end{figure}

>>>>

\begin{figure}[t!]
\begin{small}
\begin{eqnarray*}
\begin{tabular}{lc}
\textsc{[Instr]} &
W(\mbox{instr}) = \mbox{given by the spec.}
\\\\
\textsc{[Seq]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2}{W(S_1;S_2;) = w_1 + w_2}
\\\\
\textsc{[Branch]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ W(S_3) = w_3}
{W(\mbox{if}(S_1) \ S_2; \ \mbox{else} \ S_3;) = w_1 + \mbox{max}(w_2 + w_3)}
\\\\
\textsc{[LoopInf]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ \mathit{Iter}(S_1) = \infty}
{W(\mbox{while} (S_1) \ S_2;) = 10 \times (w_1 + w_2)}
\\\\
\textsc{[LoopExp]} &
\inferrule{W(S_1) = w_1 \\ W(S_2) = w_2 \\ \mathit{Iter}(S_1) = E}
{W(\mbox{while} (S_1) \ S_2;) = E \times (w_1 + w_2)}
\end{tabular}
\end{eqnarray*}
\end{small}
\caption{\label{fig:swe}Estimating the cost of tasks.}
\end{figure}

@@

Similarly, recursion bounding (Sec.~\ref{sub:ir}), although
simple, is effective in eliminating the excessive number of tasks.
Figure~\ref{fig:Optimizations} shows the effect of this optimization upon
three \textsf{BSC-Bots} benchmarks: \textsf{health}, \textsf{nqueens} and
\textsf{strassen}.
These programs were designed to illustrate the parallelization of
divide-and-conquer algorithms~\cite{Duran09}, and contain a large number of
recursive calls (see \textbf{RFC}) in Figure~\ref{fig:Optimizations}.
Pruning at higher levels of the recursion tree lets them deliver
non-negligible speedups onto the baseline programs.
Were pruning absent, then we would observe slowdowns in \textsf{health}
and \textsf{nqueens}.
This experiment shows that {\em our optimizations are effective to
improve the quality of the code that we generate}.

@@

The benchmarks used in Section~\ref{sub:performance} have been coded to
demonstrate the power of parallel systems; hence, they have
been written in a way that simplifies the discovery of parallelism in programs.
In this section, we address the following question: ``can \Taskminer" find
parallelism in general programs?
To answer it, we have applied \Taskminer{} onto the 219 C programs available in
the LLVM test suite, and have compared the annotated version
against the sequential version compiled with \textsf{clang} -O3.
\Taskminer{} annotates a benchmark if it can find symbolic bounds to every
memory access (load or store) used within a region marked as a task candidate.
Under this constraint, we have been able to discover tasks in 63 benchmarks.
Figure~\ref{fig:TM_Versatility} relates the number of tasks and the number of
instructions in the 30 largest benchmarks that we analyze.
To avoid counting multiple C files for the same benchmark,
Figure~\ref{fig:TM_Versatility} contains only benchmarks that
consists of a single file (present in LLVM's \textsf{SingleSource} folder).
As this experiment demonstrates, \Taskminer{} {\em can annotate a non-trivial
number of real-world benchmarks}.

@@

Most of the benchmarks, e.g., 27, that have been automatically annotated did not
give us speedups, because the regions marked as tasks were too small to influence
the program's runtime.
We have observed slowdowns in 17 benchmarks.
In this case, interactions between data-structures end up forcing the OpenMP
runtime to serialize execution of tasks.
Most of these slowdowns were inferior to 10\%.
In one case, \textsf{MiBench/office-stringsearch}, our program was 11x slower.
Nevertheless, we have measured speedups above 5\% in 19 benchmarks.
In \textsf{Misc/lowercase}, the speedup is above 3.5x, and in
three cases, it is above 1.5x.
Figure~\ref{fig:TM_GeneralProgs} summarizes these findings.
We emphasize that this experiment did not involve any human intervention.
Hence, {\em \Taskminer{} enables the discovery of parallelism at
zero programming cost}.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/TM_GeneralProgs}
\caption{Speedups (in number of times) obtained by \Taskminer{} when applied onto
the LLVM test suite.
The larger the bar, the better.
\textbf{LoC} stands for ``Lines of C".}
\label{fig:TM_GeneralProgs}
\end{center}
\end{figure}
